---
title: "Chapter3_GO_terms"
author: "Milena Trabert"
date: "2026-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# package install and setup

I am using this tutorial: https://bioconductor.org/packages//release/bioc/vignettes/topGO/inst/doc/topGO_manual.html


```{r}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install()
BiocManager::install("topGO")
BiocManager::install("GO.db")
```

load all packages
```{r}
library(topGO)
library(plyr)
```

read data
```{r}
setwd("/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment")

annotation_path="/Users/miltr339/work/chapter3/Cmac_functional_annot/C_mac_eggnog_diamond.emapper.annotations"
annot_header = c("query","seed_ortholog","evalue","score","eggNOG_OGs","max_annot_lvl","COG_category","Description","Preferred_name","GOs","EC","KEGG_ko","KEGG_Pathway","KEGG_Module","KEGG_Reaction","KEGG_rclass","BRITE","KEGG_TC","CAZy","BiGG_Reaction","PFAMs")
annotation = read.csv(annotation_path, sep="\t", comment.char = "#", header=FALSE, col.names = annot_header)

all_geneIDs_A_all_species_union = strsplit(readLines("all_geneIDs_A_all_species_union.txt"), ",")[[1]]
all_geneIDs_X_all_species_union = strsplit(readLines("all_geneIDs_X_all_species_union.txt"), ",")[[1]]
sig_geneIDs_A_all_species_union = strsplit(readLines("sig_geneIDs_A_all_species_union.txt"), ",")[[1]]
sig_geneIDs_X_all_species_union = strsplit(readLines("sig_geneIDs_X_all_species_union.txt"), ",")[[1]]

geneIDs_A_A_obtectus_sig = strsplit(readLines("geneIDs_A_A_obtectus_sig.txt"), ",")[[1]]
geneIDs_A_B_siliquastri_sig = strsplit(readLines("geneIDs_A_B_siliquastri_sig.txt"), ",")[[1]]
geneIDs_A_C_chinensis_sig = strsplit(readLines("geneIDs_A_C_chinensis_sig.txt"), ",")[[1]]
geneIDs_X_A_obtectus_sig = strsplit(readLines("geneIDs_X_A_obtectus_sig.txt"), ",")[[1]]
geneIDs_X_B_siliquastri_sig = strsplit(readLines("geneIDs_X_B_siliquastri_sig.txt"), ",")[[1]]
geneIDs_X_C_chinensis_sig = strsplit(readLines("geneIDs_X_C_chinensis_sig.txt"), ",")[[1]]

geneIDs_A_A_obtectus = read.csv("geneIDs_A_A_obtectus_all.txt", sep=",", header=TRUE)
geneIDs_A_B_siliquastri = read.csv("geneIDs_A_B_siliquastri_all.txt", sep=",", header=TRUE)
geneIDs_A_C_chinensis = read.csv("geneIDs_A_C_chinensis_all.txt", sep=",", header=TRUE)
geneIDs_X_A_obtectus = read.csv("geneIDs_X_A_obtectus_all.txt", sep=",", header=TRUE)
geneIDs_X_B_siliquastri = read.csv("geneIDs_X_B_siliquastri_all.txt", sep=",", header=TRUE)
geneIDs_X_C_chinensis = read.csv("geneIDs_X_C_chinensis_all.txt", sep=",", header=TRUE)
```

make into topGO data structure: 
1. start with making the enrichment geneIDs lists into the required named vectors (since I have binary categories and not p-values as expected, the data frames have kind of pseudo-p-values that are 0.0 if the gene is positively selected and 1 otherwise)
```{r}
make_gene_list <- function(input_df){
  geneList <- factor(as.integer(input_df$pos_sel))
  names(geneList) <- input_df$geneID
  return(geneList)
}

A_A_obtectus_enrichment_list <- make_gene_list(geneIDs_A_A_obtectus)
A_B_siliquastri_enrichment_list <- make_gene_list(geneIDs_A_B_siliquastri)
A_C_chinensis_enrichment_list <- make_gene_list(geneIDs_A_C_chinensis)
X_A_obtectus_enrichment_list <- make_gene_list(geneIDs_X_A_obtectus)
X_B_siliquastri_enrichment_list <- make_gene_list(geneIDs_X_B_siliquastri)
X_C_chinensis_enrichment_list <- make_gene_list(geneIDs_X_C_chinensis)

make_contrasts_for_AvsX <- function(list_A, list_X){
  all_geneIDs <- append(list_A,list_X)
  AX_presence <- rep(FALSE, length(all_geneIDs))
  for (i in 1:length(all_geneIDs)){
    if (all_geneIDs[i] %in% list_X){
      AX_presence[i]<-TRUE
    }
  }
  geneList <- factor(as.integer(AX_presence))
  names(geneList) <- all_geneIDs
  return(geneList)
}

all_AX_contrast <- make_contrasts_for_AvsX(list_A = all_geneIDs_A_all_species_union, list_X = all_geneIDs_X_all_species_union)
sig_AX_contrast <- make_contrasts_for_AvsX(list_A = sig_geneIDs_A_all_species_union, list_X = sig_geneIDs_X_all_species_union)
```

prepare eggnog annotation file
```{r}
valid <- !is.na(annotation$GOs) & annotation$GOs != "-"
gene2GO <- strsplit(annotation$GOs[valid], ",")
names(gene2GO) <- annotation$query[valid]
```


make GO enrichment object for each contrast
```{r}

GO_enrichment_analysis <- function(geneList, annotation_df){
  # prepare input data
  ## downsample annotation to only include genes in the current test
  count(geneList)
  gene2GO <- annotation_df[names(annotation_df) %in% names(geneList)]
  ## create topGO object
  # ontology = BP means that it's all biological process level GOs
  GOdata <- new("topGOdata",ontology = "BP",allGenes = geneList,annot = annFUN.gene2GO,gene2GO = gene2GO)
  
  # run enrichment test
  result <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
  ## about the "elim" algorithm from the documentation: "Here one computes the p-value of a GO term conditioned on the neighbouring terms. The tests are therefore not independent and the multiple testing theory does not directly apply. We like to interpret the p-values returned by these methods as corrected or not affected by multiple testing."
  
  allRes <- GenTable(GOdata,classicFisher = result,orderBy = "classicFisher",topNodes = 20)
  return(allRes)
}

A_A_obtectus_enriched_GOs <- GO_enrichment_analysis(geneList = A_A_obtectus_enrichment_list, annotation_df = gene2GO)
A_B_siliquastri_enriched_GOs <- GO_enrichment_analysis(geneList = A_B_siliquastri_enrichment_list, annotation_df = gene2GO)
A_C_chinensis_enriched_GOs <- GO_enrichment_analysis(geneList = A_C_chinensis_enrichment_list, annotation_df = gene2GO)
X_A_obtectus_enriched_GOs <- GO_enrichment_analysis(geneList = X_A_obtectus_enrichment_list, annotation_df = gene2GO)
X_B_siliquastri_enriched_GOs <- GO_enrichment_analysis(geneList = X_B_siliquastri_enrichment_list, annotation_df = gene2GO)
X_C_chinensis_enriched_GOs <- GO_enrichment_analysis(geneList = X_C_chinensis_enrichment_list, annotation_df = gene2GO)

write.csv(A_A_obtectus_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/A_A_obtectus_enriched_GOs.csv", row.names=TRUE)
write.csv(A_B_siliquastri_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/A_B_siliquastri_enriched_GOs.csv", row.names=TRUE)
write.csv(A_C_chinensis_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/A_C_chinensis_enriched_GOs.csv", row.names=TRUE)
write.csv(X_A_obtectus_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/X_A_obtectus_enriched_GOs.csv", row.names=TRUE)
write.csv(X_B_siliquastri_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/X_B_siliquastri_enriched_GOs.csv", row.names=TRUE)
write.csv(X_C_chinensis_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/X_C_chinensis_enriched_GOs.csv", row.names=TRUE)
```








