---
title: "Chapter3_GO_terms"
author: "Milena Trabert"
date: "2026-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Enrichment with TopGo

I am using this tutorial: https://bioconductor.org/packages//release/bioc/vignettes/topGO/inst/doc/topGO_manual.html

## setup

install
```{r}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install()
BiocManager::install("topGO")
BiocManager::install("GO.db")
BiocManager::install("ViSEAGO")
```

load all packages
```{r}
library(topGO)
library(plyr)
```

## enrichment analysis

I do everything here but we only use the A/X general and C. chinensis species-specific comparisons.
read data
```{r}
setwd("/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment")

annotation_path="/Users/miltr339/work/chapter3/Cmac_functional_annot/C_mac_eggnog_diamond.emapper.annotations"
annot_header = c("query","seed_ortholog","evalue","score","eggNOG_OGs","max_annot_lvl","COG_category","Description","Preferred_name","GOs","EC","KEGG_ko","KEGG_Pathway","KEGG_Module","KEGG_Reaction","KEGG_rclass","BRITE","KEGG_TC","CAZy","BiGG_Reaction","PFAMs")
annotation = read.csv(annotation_path, sep="\t", comment.char = "#", header=FALSE, col.names = annot_header)

### files for pairwise site model comparisons --> outdated!
# all_geneIDs_A_all_species_union = strsplit(readLines("all_geneIDs_A_all_species_union.txt"), ",")[[1]]
# all_geneIDs_X_all_species_union = strsplit(readLines("all_geneIDs_X_all_species_union.txt"), ",")[[1]]
# sig_geneIDs_A_all_species_union = strsplit(readLines("sig_geneIDs_A_all_species_union.txt"), ",")[[1]]
# sig_geneIDs_X_all_species_union = strsplit(readLines("sig_geneIDs_X_all_species_union.txt"), ",")[[1]]

### bruchini-site model files 
nonsig_geneIDs_A = strsplit(readLines("bruchini_site_model_beta_Cmac_geneIDs_all_A.txt"), ",")[[1]]
nonsig_geneIDs_X = strsplit(readLines("bruchini_site_model_beta_Cmac_geneIDs_all_X.txt"), ",")[[1]]
sig_geneIDs_A = strsplit(readLines("bruchini_site_model_beta_Cmac_geneIDs_sig_A.txt"), ",")[[1]]
sig_geneIDs_X = strsplit(readLines("bruchini_site_model_beta_Cmac_geneIDs_sig_X.txt"), ",")[[1]]

nonsig_geneIDs = append(nonsig_geneIDs_A, nonsig_geneIDs_X)
sig_geneIDs = append(sig_geneIDs_A, sig_geneIDs_X)
A_geneIDs = append(nonsig_geneIDs_A, sig_geneIDs_A)
X_geneIDs = append(nonsig_geneIDs_X, sig_geneIDs_X)

```

make into topGO data structure: 
start with making the enrichment geneIDs lists into the required named vectors (since I have binary categories and not p-values as expected, the data frames have values that are 1 if the gene is positively selected and 0 otherwise)
```{r}

make_contrasts_for_AvsX <- function(background_excl, foreground){
  all_geneIDs <- append(background_excl,foreground)
  union_set <- rep(FALSE, length(all_geneIDs))
  for (i in 1:length(all_geneIDs)){
    if (all_geneIDs[i] %in% foreground){
      union_set[i]<-TRUE
    }
  }
  geneList <- factor(as.integer(union_set))
  names(geneList) <- all_geneIDs
  return(geneList)
}

AX_contrast <- make_contrasts_for_AvsX(background_excl = A_geneIDs, foreground = X_geneIDs)
nonsig_AX_contrast <- make_contrasts_for_AvsX(background_excl = nonsig_geneIDs_A, foreground = nonsig_geneIDs_X)
sig_AX_contrast <- make_contrasts_for_AvsX(background_excl = sig_geneIDs_A, foreground = sig_geneIDs_X)
A_sig_contrast <- make_contrasts_for_AvsX(background_excl = nonsig_geneIDs_A, foreground = sig_geneIDs_A)
X_sig_contrast <- make_contrasts_for_AvsX(background_excl = nonsig_geneIDs_X, foreground = sig_geneIDs_X)
all_sig_contrast <- make_contrasts_for_AvsX(background_excl = nonsig_geneIDs, foreground = sig_geneIDs)
```

prepare eggnog annotation file
```{r}
valid <- !is.na(annotation$GOs) & annotation$GOs != "-"
gene2GO <- strsplit(annotation$GOs[valid], ",")
names(gene2GO) <- annotation$query[valid]
```

## run enrichment

make GO enrichment object for each contrast
```{r}

GO_enrichment_analysis <- function(geneList, annotation_df){
  # prepare input data
  ## downsample annotation to only include genes in the current test
  count(geneList)
  gene2GO <- annotation_df[names(annotation_df) %in% names(geneList)]
  ## create topGO object
  # ontology = BP means that it's all biological process level GOs
  GOdata <- new("topGOdata",ontology = "BP",allGenes = geneList,annot = annFUN.gene2GO,gene2GO = gene2GO)
  
  # run enrichment test
  result <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
  ## about the "elim" algorithm from the documentation: "Here one computes the p-value of a GO term conditioned on the neighbouring terms. The tests are therefore not independent and the multiple testing theory does not directly apply. We like to interpret the p-values returned by these methods as corrected or not affected by multiple testing."
  
  ## check for presence of the dosage compensation GO-term
  annotated_genes <- genesInTerm(GOdata, "GO:0009047")[["GO:0009047"]]
  print(annotated_genes)

  allRes <- GenTable(GOdata,classicFisher = result,orderBy = "classicFisher",topNodes = 1000)
  return(allRes)
}

### A to X comparison
all_AX_enriched_GOs <- GO_enrichment_analysis(geneList = AX_contrast, annotation_df = gene2GO)
nonsig_AX_enriched_GOs <- GO_enrichment_analysis(geneList = nonsig_AX_contrast, annotation_df = gene2GO)
sig_AX_enriched_GOs <- GO_enrichment_analysis(geneList = sig_AX_contrast, annotation_df = gene2GO)

### sig to nonsig comparison
A_sig_enriched_GOs <- GO_enrichment_analysis(geneList = A_sig_contrast, annotation_df = gene2GO)
X_sig_enriched_GOs <- GO_enrichment_analysis(geneList = X_sig_contrast, annotation_df = gene2GO)
all_sig_enriched_GOs <- GO_enrichment_analysis(geneList = all_sig_contrast, annotation_df = gene2GO)

write.csv(all_AX_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/all_AX_enriched_GOs_revision.csv", row.names=TRUE)
write.csv(nonsig_AX_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/nonsig_AX_enriched_GOs_revision.csv", row.names=TRUE)
write.csv(sig_AX_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/sig_AX_enriched_GOs_revision.csv", row.names=TRUE)
write.csv(A_sig_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/A_sig_enriched_GOs_revision.csv", row.names=TRUE)
write.csv(X_sig_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/X_sig_enriched_GOs_revision.csv", row.names=TRUE)
write.csv(all_sig_enriched_GOs, "/Users/miltr339/work/PhD_code/PhD_chapter3/data/GO_enrichment/all_sig_enriched_GOs_revision.csv", row.names=TRUE)
```

# Try semantic clustering with Viseago

This did not work out but I will keep it for posterity just in case

## load libraries
```{r}
library(ggplot2)
library(GO.db)
library(ViSEAGO)
```

## read data

background annotation
```{r}
filter_GO_terms_background <- function(GOfil){
  # GOfil is a dataframe
  # make go terms column to list
    GOfil$GOs <- lapply(GOfil$GOs, function(x) {
    if (x == "-") {
      x
    } else {
      strsplit(x, ",")[[1]]
    }
  })
  org_n<-nrow(GOfil)
  genes_before_filtering <- length(unique(GOfil$query))
  GOfil<-GOfil[which(GOfil$GOs %in% keys(GO.db)),]
  new_n<-nrow(GOfil)
  genes_after_filtering <- length(unique(GOfil$query))
  print(filename)
  print("number of genes lost:")
  print(genes_before_filtering - genes_after_filtering)
  # Looks like one gene was lost
  print("number of GO-terms lost:")
  print(org_n-new_n)
  print("------------------------------")
  edited_filename <- sub("\\.tsv$", "_ViSEAGO_filtered", filename)
  write.table(GOfil,edited_filename,sep="\t",col.names=TRUE,row.names=FALSE,quote=FALSE)
  return(edited_filename)
}

annotation_path="/Users/miltr339/work/chapter3/Cmac_functional_annot/C_mac_eggnog_diamond.emapper.annotations.tsv"
annot_header = c("query","seed_ortholog","evalue","score","eggNOG_OGs","max_annot_lvl","COG_category","Description","Preferred_name","GOs","EC","KEGG_ko","KEGG_Pathway","KEGG_Module","KEGG_Reaction","KEGG_rclass","BRITE","KEGG_TC","CAZy","BiGG_Reaction","PFAMs")
GOfil = read.table(annotation_path, sep="\t", comment.char = "#", header=FALSE, col.names = annot_header)
filter_GO_terms_background(GOfil = annotation)

```






